### 2012-07-07 src/simulation_schema.tex

\documentclass{article}

\usepackage[latin1]{inputenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{pst-eps}
\begin{document}
\pagestyle{empty}

% Define block styles
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    text width=8em, text centered, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, text centered,text width=7em, 
    minimum height=2em]
\tikzstyle{dummy} = [draw, ellipse, text centered,text width=7em, 
    minimum height=2em]

\begin{TeXtoEPS}
\begin{tikzpicture}[auto]
    % Place nodes
    \node [cloud] (gt) {$X$ \\ (ground-truth)};
    \node [block, right of=gt, node distance = 16em] (system) {SAR \\ Processor};
    \node [right of=system, node distance = 16em] (dummy1) {};
    \node [cloud, below of=dummy1, node distance = 6em] (observed) {$\tilde{X}$ \\ (observed)};
    \node [below of=observed, node distance = 6em] (dummy2) {};
    \node [block, left of=dummy2, node distance = 16em] (filter) {SAR \\ Speckle Filter};
    \node [cloud, left of=filter, node distance = 16em] (estimated) {$\hat{X}$ \\ (estimated)};
    % Draw edges
    \path [line] (gt) -- (system);
    \path [line] (system) -| (observed);
    \path [line] (observed) |- (filter);
    \path [line] (filter) -- (estimated);
\end{tikzpicture}
\end{TeXtoEPS}

\end{document}